---
title: "Data Cleaning and Manipulation"
output: pdf_document
author: Anastasija Mensikova
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
library(dplyr)
library(tidyr)

# To call stats filter -- stats::filter()
```


## Reformat the catch data

* Remove "all" column
* Remove "notes" column
* Create a "species" column
    - Move from wide to long
* General QA


[Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/view/df35b.304.2)

```{r}
catch_original = read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", stringsAsFactors = FALSE)

head(catch_original)
```


### Remove the "all" and "notesRegCode" columns using "select"

```{r}
#Cmd + shift + m: pipe operator shortcut ( %>% )

# catch_long = catch_original %>% select(-All, -notesRegCode) %>% ... --> this will be the same

catch_long = catch_original %>% select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>% 
                                gather(key = "species", value = "catch", -Region, -Year)

head(catch_long)
```

```{r}
catch_wide = catch_long %>% spread(key = species, value = catch)

head(catch_wide)
```


### Clean up our data

* Rename catch to catch_thousands
* Change "catch" column to numeric
* Create a new "catch" column in units num. of fish


```{r}
catch_clean = catch_long %>% rename(catch_thousands = catch) %>%
                             mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>% 
                             mutate(catch_thousands = as.numeric(catch_thousands)) %>% 
                             mutate(catch = catch_thousands * 1000) %>% 
                             select(-catch_thousands)

head(catch_clean)
```


```{r, eval = F, echo = F}
# Look for where as.numeric failed (NAs introduced by coercion)
test = as.numeric(catch_long$catch)
i = which(is.na(test) == TRUE)
catch_long[i,]
```


### Split - Apply - Combine

* Calculate mean catch by species


```{r}
species_mean = catch_clean %>% group_by(species, Region) %>% summarise(catch_mean = mean(catch), num_obs = n())

head(species_mean)
```


```{r}
year_mean = catch_clean %>% group_by(Year) %>% summarise(catch_mean = mean(catch)) %>% arrange(-catch_mean) %>% 
            filter(Year >= 1990)

year_mean
```



## Join the region definitions

```{r}
region_defs = read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", stringsAsFactors = FALSE) %>% select(code, mgmtArea) %>% rename(Region = code)
head(region_defs)
```


```{r}
catch_joined = left_join(catch_clean, region_defs, by = c("Region"))

head(catch_joined)
```


```{r}
dates_df = data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

dates_split = dates_df %>% separate(date, c("month", "day", "year"), "/", remove = F)

dates_split
```


```{r}
dates_unite = dates_split %>% unite(col = date_iso, year, month, day, sep = "-")

head(dates_unite)
```



## Miscellaneous functions

