---
title: "Data Visualisation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(leaflet)
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
```


## Reformat the data for plotting

Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.

```{r}
esc = read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", stringsAsFactors = F)

head(esc)
```


* Calculate annual escapement by species and region
* Filter for just pacific salmon (Sockeye, Chinook, Coho, Pink, Chum)


```{r}
annual_esc = esc %>% separate(sampleDate, c("year", "month", "day"), "-", remove = F) %>% 
                     # corece year to numeric type
                     mutate(year = as.numeric(year)) %>% 
                     group_by(SASAP.Region, Species, year) %>% 
                     summarise(count = sum(DailyCount))

head(annual_esc)
```


```{r}
annual_esc = annual_esc %>% filter(Species %in% c("Sockeye", "Chinook", "Coho", "Pink", "Chum"))

head(annual_esc)
```


## Generate some static plots

```{r}
ggplot(annual_esc, aes(x = Species, y = count, fill = SASAP.Region)) + 
  geom_col()
```



```{r}
my_theme = theme_test() +
  theme(legend.position = "bottom", legend.title = element_blank())

ggplot(filter(annual_esc, SASAP.Region == "Kodiak"), aes(x = year, y = count, color = Species)) +
  geom_line() +
  geom_point() +
  ylab("Escapement (# of fish)") +
  xlab("Year") +
  ggtitle("Kodiak Salmon Escapement") +
  my_theme
```



## Generate an interactive map

```{r}

```



